Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2016-01-21 body/conclusion.tex
added missing figure on feasibility
small fixes in revised sections

In this paper we consider acyclic contact planning in cluttered environments, formulated as two sub problems that we address sequentially.
The first problem is \Pa, that consists in computing a guide path for the root of the robot that can be extended; the second problem is \Pb, that is the computation of a discrete sequence of contacts along the root path.
Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. We introduced a low-dimensional space $C_{reach}$ that can be mapped 
into the contact sub-manifold of the robot, approximated, and efficiently sampled by our Reachability-Based planner.
Our contribution to \Pb is a pragmatic contact generation scheme that can take into
account criteria to enforce interesting properties on the generated contacts (such as robustness, energy efficiency or naturalness).

Aside from the theoretical contributions, our results demonstrate that our method allows a very interesting compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \gls{reachc}, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework is really efficient in addressing both \Pa and \Pb. This results in computation costs close to real-time in
known static environments.
%
\textbf{Regarding the quality of the paths}, a parametrization of the \gls{reachc} allows us to compute
relevant paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the compromise criterion used in computing the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already obtained some successes for some of the computed sequences~\citep{Carpentier2016}, but we admit that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible.
Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
at the center of mass of the robot.
 We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
Lastly, we aim at going beyond static equilibrium and to perform kynodynamic planning in the configuration space. We believe that the most promising direction in this regard is to integrate
the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these three issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.

>>>>

In this paper we consider acyclic contact planning in cluttered environments, formulated as two sub problems that we address sequentially.
The first problem is \Pa, that consists in computing a guide path for the root of the robot that can be extended; the second problem is $\mathcal{P}_2$, that is the computation of a discrete sequence of contacts along the root path.
Our contribution to \Pa is a generic characterization of the properties that the guide path must satisfy, in particular to enforce the completeness of the acyclic contact planner. We introduced a low-dimensional space $C_{reach}$ that can be mapped 
into the contact sub-manifold of the robot, approximated, and efficiently sampled by our Reachability-Based planner.
Our contribution to \Pb is a pragmatic contact generation scheme that can take into
account criteria to enforce interesting properties on the generated contacts (such as robustness, energy efficiency or naturalness).

Aside from the theoretical contributions, our results demonstrate that our method allows a very interesting compromise between three 
criteria that are hard to conciliate: generality, performance, and quality of the solution, making it the first acyclic contact
planner compatible with \gls{interactive} applications.
%
\textbf{Regarding generality}, the \gls{reachc}, coupled with an approach based on limb decomposition, 
allows the method to address arbitrary multiped robots. The only pre-requisite is the specification 
of the volumes $W^0$.
%
\textbf{Regarding performance}, our framework is really efficient in addressing both \Pa and $\mathcal{P}_2$. This results in computation costs close to real-time in
known static environments.
%
\textbf{Regarding the quality of the paths}, a parametrization of the \gls{reachc} allows us to compute
relevant paths in all the presented scenarios, with low rejection rates.
As for \cite{Bouyarmane2009}, failures can still occur, due to the compromise criterion used in computing the guide path.
The low computational burden of our framework however allows for fast re-planning in case of failure.

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already obtained some successes for some of the computed sequences~\citep{Carpentier2016}, but we admit that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible.
Our current direction is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
at the center of mass of the robot.
 We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
Lastly, we aim at going beyond static equilibrium and to perform kinodynamic planning in the configuration space. We believe that the most promising direction in this regard is to integrate
the notion of Admissible Velocity Propagation in our current work \citep{DBLP:conf/rss/PhamCN13}.
Addressing these three issues is essential to bridge the gap between the planning and control aspects of multiped locomotion.

### 2016-01-21 body/intro.tex
small fixes in revised sections

To address \Pa, we first use the \textit{reachability condition} to plan the guide path. This step is fast because the planning happens
in a low-dimensional space, and does not require explicit contact computations.

@@

\subsection{Paper contribution and organization}
We propose contributions to both problems \Pa and \Pb~.
\begin{itemize}
\item A low-dimensional, efficient sampling-based approach to plan guide paths.
\item A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.
\item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
more efficient with respect to the task.
\item Statistical tests that empirically demonstrate the validity of our approach in the considered scenarios. 
\end{itemize}

@@

In Section~\ref{overview}, we present an overview of our method. Section~\ref{rbprm} and Section~\ref{sec:contact} present respectively our answer to \Pa and \Pb. In Section~\ref{sec:heuristics}, we present heuristics used for the selection of a contact configuration. Finally, in Section 6 we propose a complete experimental validation of the planner with three very different kinematic chains (the HRP-2 and HyQ robots, and a three-finger manipulator) in various scenarios.
%that complete those presented in the conference paper.

>>>>

To address $\mathcal{P}_1$, we first use the \textit{reachability condition} to plan the guide path. This step is fast because the planning happens
in a low-dimensional space, and does not require explicit contact computations.

@@

\subsection{Paper contribution and organization}
We propose contributions to both problems \Pa and $\mathcal{P}_2$.
\begin{itemize}
\item A low-dimensional, efficient sampling-based approach to plan guide paths.
\item A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.
\item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
more efficient with respect to the task.
\item Statistical tests that empirically demonstrate the validity of our approach in the considered scenarios. 
\end{itemize}

@@

In Section~\ref{overview}, we present an overview of our method. Section~\ref{rbprm} and Section~\ref{sec:contact} present respectively our answer to \Pa and $\mathcal{P}_2$. In Section~\ref{sec:heuristics}, we present heuristics used for the selection of a contact configuration. Finally, in Section 6 we propose a complete experimental validation of the planner with three very different kinematic chains (the HRP-2 and HyQ robots, and a three-finger manipulator) in various scenarios.
%that complete those presented in the conference paper.

### 2016-01-21 body/overview.tex
small fixes in revised sections

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of \Pa, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is too computationally expensive because it requires the projection of the whole-body configuration into a contact posture.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$). 
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

>>>>

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). The root of the robot is indicated with a black cross. To compute the reachable workspace, the point on the ankle indicated by a green cross was used. $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of \Pa, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is too computationally expensive because it requires the projection of the whole-body configuration into a contact posture.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$). Abusively, for simplicity,
we refer to the root as the kinematic chain described by all these DOFS.
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the root of the robot is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated according to the root configuration $\mathbf{q}^{0}$.

### 2016-01-21 body/results.tex
small fixes in revised sections

\subsubsection{On the choice of heuristics} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we now prefer to use our manipulability based heuristic for the legs of the robot, but we still
use EFORT for the arm, that results in lesser contact repositionning.

>>>>

\subsubsection{On the choice of heuristics} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we now prefer to use our manipulability based heuristic for the legs of the robot, but we still
use EFORT for the arm, that results in less contact repositionning.
